\htmlhr
\chapter{Installation and app setup\label{installation}}

\section {Requirements\label{sec:requirements}}
\paragraph{Java 7}
\begin{itemize}
 \item  \<.../jdk1.7.0/bin> must be on your path.
 \item \<JAVA\_HOME> should be set to \<.../jdk1.7.0>.
\end{itemize}

\paragraph{Android SDK}
\begin{itemize}
 \item Android API version 15 or later
\end{itemize}
\begin{enumerate}
 \item Install the Android SDK to some directory.
 \item Set \<ANDROID\_HOME> to the directory where you installed the
   Android SDK.
 \item Download the \<android-15> target by running \<\$ANDROID\_HOME/tools/android>
\end{enumerate}

%If using Eclipse, go to
%\<Help $\rightarrow$ Install New Software>
%and install the Android ADT Plugin (\url{https://dl-ssl.google.com/android/eclipse}).

\section{Install SPARTA}

\begin{enumerate}
\item Download the SPARTA toolset: \url{https://github.com/typetools/sparta/releases/download/sparta-1.0.2/sparta-toolset-1.0.2.tgz}

\item Unpack the archive

\item Set \<SPARTA\_CODE> to the directory where you unpacked the toolset.
\end{enumerate}

See Section~\ref{install-source} for instructions on how to install SPARTA from source.

The SPARTA toolset can be used with apps that build using Ant (Section~\ref{install-ant})
or using the Android plugin for Gradle (Section~\ref{install-gradle}).

\subsection{Using SPARTA with the Android plugin for Gradle\label{install-gradle}}
\begin{enumerate}
\item Use the Maven Central repository:

\begin{Verbatim}
repositories {
  ... existing repositories...
  mavenCentral()
}
\end{Verbatim}

\item Add dependency configurations for the annotated JDK, \<checker.jar>, and the Type Annotations compiler:

\begin{small}
\begin{Verbatim}
configurations {
    ... existing configurations ...
    checkerFrameworkJavac {
        description = 'a customization of the Open JDK javac compiler with additional support for type annotations'
    }
    checkerFramework {
       description = 'The Checker Framework: custom pluggable types for Java'
    }
    sparta {
       description =  `Static Program Analysis for Reliable Trusted Apps`
    }
}
\end{Verbatim}
\end{small}

\item Declare the Checker Framework and SPARTA dependencies:

\begin{small}
\begin{Verbatim}
dependencies {
    ... existing dependencies...
    ext.checkerFrameworkVersion = '1.9.11'
    checkerFrameworkJavac "org.checkerframework:compiler:${checkerFrameworkVersion}"
    checkerFramework "org.checkerframework:checker:${checkerFrameworkVersion}"
    compile "org.checkerframework:checker-qual:${checkerFrameworkVersion}"

    sparta fileTree(dir: "$System.env.SPARTA_CODE", include: ['*.jar'])
    if (typecheck) {
        compile fileTree(dir: "$System.env.SPARTA_CODE", include: ['*.jar'])
    }

}
\end{Verbatim}
\end{small}

\item Direct all tasks of type \<JavaCompile> to use the desired checkers, when the \<typecheck> project property is set:

\begin{small}
\begin{Verbatim}
def typecheck = project.properties['typecheck'] ?: false
allprojects {
if (typecheck) {
    gradle.projectsEvaluated {
        tasks.withType(JavaCompile).all { JavaCompile compile ->
        compile.options.compilerArgs = [
                '-processor', 'sparta.checkers.FlowChecker',
                '-processorpath', "${configurations.sparta.asPath}:${configurations.checkerFramework.asPath}",

        ]
            compile.options.compilerArgs += ['-source', '7', '-target', '7']
            options.bootClasspath = "${configurations.checkerFrameworkJavac.asPath}:" + System.getProperty("sun.boot.class.path") + ":" + options.bootClasspath

            options.fork = true
            options.forkOptions.jvmArgs += ["-Xbootclasspath/p:${configurations.checkerFrameworkJavac.asPath}"]
        }
    }
}
}

\end{Verbatim}
\end{small}

\item Run the checkers

\begin{Verbatim}
gradle compileReleaseJavaWithJavac -Ptypecheck=true
\end{Verbatim}

\end{enumerate}



\subsection{Using SPARTA with Ant\label{install-ant}}

\begin{enumerate}

\item Install the Checker Framework by following the installation instructions in the manual:
\url{http://types.cs.washington.edu/checker-framework/current/checker-framework-manual.html#installation}

\item
Update the app configuration by running the following command in the main directory of the
app.
\begin{Verbatim}
$ANDROID_HOME/tools/android update project --path .
\end{Verbatim}

\item
Copy the file \<sparta.jar> generated in Section~\ref{sec:install} to the
directory \<libs/> in the main directory of the app. If that folder doesn't
exist it must be created.

\item Build the app

\begin{Verbatim}
ant release
\end{Verbatim}
If the app does not build with the above command, then the SPARTA tool set will
not be able to analyze the app.  Below are two common compilation issues and
solutions.
\begin{itemize}
\item
Most Android apps will rely on an auto-generated \<R.java> file
in the \</gen> directory of the project. This will only be generated
if there are no errors in the project. There may be errors in the
resources (\<.../res> directory) that could cause \<R.java> to not be
generated.

\item
Additionally, if the app depends on an external \<.jar> file, it will compile
in Eclipse but not with Ant. All dependent \<.jar> files must be in the
\<libs/> directory.
\end{itemize}

\item
Add the SPARTA build targets to the end of the \<build.xml>
 file, just above \verb|</project>|.


\begin{Verbatim}
    <property name="checker-framework" value="${env.CHECKERFRAMEWORK}"/>
    <property name="sparta-code" value="${env.SPARTA_CODE}"/>

    <dirname property="checker-framework_dir" file="${checker-framework}/checker-framework" />
    <dirname property="sparta-code_dir" file="${sparta-code}/sparta-code" />

    <import file="${sparta-code_dir}/build.include.xml" />
    <property name="flowPolicy" value="flow-policy"/>
    <property name="componentMap" value="component-map"/>
\end{Verbatim}

\end{enumerate}

\section{Installing SPARTA from source\label{install-source}}

\begin{enumerate}

\item
  Obtain the source code for the SPARTA tools, either from its version
  control repository or from a packaged release.

\begin{itemize}
\item
  To obtain from the version control repository, run
\begin{Verbatim}
git clone git@github.com:typetools/sparta.git
\end{Verbatim}
\item

The source code is also available as a part of the release:
  download the SPARTA release from
  \url{https://types.cs.washington.edu/sparta/release/}.

  Then, unpack the archive.
\end{itemize}

%
%Rationale:  When working with Android, a developer must ``update a project'' to
%properly set the path to the Android SDK\@.  For more details about
%updating an Android project, see
%\url{http://developer.android.com/tools/projects/projects-cmdline.html#UpdatingAProject}.


\item Build the SPARTA tools by compiling the source code:
\begin{alltt}
ant jar
\end{alltt}

\item
As a sanity check of the installation, run

\begin{alltt}
ant all-tests
\end{alltt}

You should see ``\<BUILD SUCCESSFUL>'' at the end.


\end{enumerate}

\section{Using Eclipse to analyze apps}
To use Eclipse to look at and build the code, perform these
steps:
\begin{itemize}
\item
Import the projects the app.      \<Import $\rightarrow$ Existing Android Code
Into Workspace>

    \item
    Make sure
    \<Project Properties $\rightarrow$ Android $\rightarrow$ Android
    version \#> is checked

    \item
    Check that
    \<Project Properties $\rightarrow$ Java Build Path $\rightarrow$
    Libraries $\rightarrow$ Android version \#> appears

    \item
    Add sparta.jar to the apps build path

    \item Right click on the build.xml file and select   \<Run as $\rightarrow$
    External Tools Configurations...>. In the \<Main> tab add \<check-flow> to the
    Arguments box.  In the \<Environment> tab, add the \<CHECKERFRAMEWORK> and
    \<SPARTA\_CODE> variables.

  \end{itemize}



%%% Local Variables:
%%% mode: latex
%%% TeX-master: "manual"
%%% TeX-command-default: "PDF"
%%% End:

%  LocalWords:  hg SDK ADT Plugin MercurialEclipse gson json sparta jsr308
%  LocalWords:  Dandroid langtools javaparser app TODO xml env dirname dir
%  LocalWords:  basename
